analysis

### 2020-06-03 analysis.Rmd
working on HGAM analysis

### Prepare environment
```{r setup, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(lme4)
library(lmerTest)
library(tidyverse)
library(lubridate)
library(GGally)
```

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      filter(freq > 1) %>% # remove singletons
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species specificity")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      #filter(freq > 1) %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher", index = c("web asymmetry", "weighted NODF", "niche overlap", "C score")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

# network level version 2
rare_net_nt2 <- function(net, n, iter) {

@@

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      select(site, date, phase, year) %>%
      unique() %>%
      group_by(site, phase, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      #filter(freq > 1) %>%
      unite(webID, c(site, phase, year), sep = "_") %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher",
          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(id = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      select(site, date, phase, year) %>%
      unique() %>%
      group_by(site, phase, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      #filter(freq > 1) %>%
      unite(webID, c(site, phase, year), sep = "_") %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "id") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

### Colony phase breakpoint analysis
```{r}
castes <- read_csv("./processed_data/network.csv") %>%
  group_by(site, date, year, caste) %>%
  summarize(caste.count = n()) %>%
  spread(caste, caste.count) %>%
  replace(is.na(.), 0) %>%
  mutate(queen = oldqueen + queen,
         reproductive = male + youngqueen,
         q.prop = queen/(queen+worker+reproductive),
         r.prop = reproductive/(queen+worker+reproductive)) %>%
  select(-oldqueen, -youngqueen, -male) %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(site_data, by = c("site")) %>%
  arrange(gdd.cum) %>%
  mutate(yday = yday(date))

breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.75, t2 = 0.1)
breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.75, t2 = 0.1)
breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.75, t2 = 0.1)

breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.1)
breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.1)
breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.1)

@@

### Plot phase breakpoints
```{r}
ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
  geom_point() +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("Julian day") +
  ylab("Proportional abundance")

ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
  geom_point() +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("GDD") +
  ylab("Proportional abundance")
```

@@

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  #semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)
```
### BB and floral turnover ~ elevation
```{r}
bb_range <- net %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

@@

ggplot(filter(fl_survey_elev, sites > 0), 
       aes(reorder(plant.sp, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_bw() +
  theme(axis.text.x = element_blank())

ggplot(filter(fl_net_elev, sites > 0), 
       aes(reorder(plant.sp, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_bw() +
  theme(axis.text.x = element_blank())

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 3, color = "gray60") +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_bw()
```

### BB abundance, diversity, and packing
```{r eval=FALSE, echo=FALSE}
### Per-transect Bombus density
bb_abund <- net %>%
  group_by(site, date, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.abund = n())

### Per-transect Bombus richness
bb_rich <- net %>%
  group_by(site, date, gdd.cum, year, elev.class2,
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.rich = length(unique(bb.sp)))

### Per-transect floral abund
fl_abund <- survey %>%
  group_by(site, date, gdd.cum, year, break1.gdd, break2.gdd) %>%
  summarize(fl.abund = sum(flower.cover),
            log.fl.abund = log(sum(flower.cover)))

### Per-transect floral richness
fl_rich <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, date, gdd.cum, year, break1.gdd, break2.gdd) %>%
  summarize(fl.rich = n())

### Join BB density and floral cover
abund_div <- bb_abund %>%
  full_join(bb_rich) %>%
  full_join(fl_abund) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
  full_join(fl_rich) %>%
  na.omit() %>%
  ungroup %>%
  mutate(fl.per.bb = fl.abund/bb.abund,
         site = factor(site),
         yday = yday(date),
         week = week(date),
         month = month(date)) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")),
         total.rich = bb.rich + fl.rich,
         web.asymmetry = (bb.rich - fl.rich) / (bb.rich + fl.rich))

@@

@@

@@

### Web asymmetry ~ gdd
ggplot(abund_div, aes(gdd.cum, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

### Web asymmetry ~ yday
ggplot(abund_div, aes(yday, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

@@

### Web asymmetry ~ Per-capita floral resource abundance
ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")
```

### Phase x elev network analysis
```{r}
### Prepare pxe abundance and diversity table
pxe_abund_div <- abund_div %>%
  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
  group_by(site, phase, year, ad.metric) %>%
  summarize(ad.value = mean(ad.value))

ggplot(filter(pxe_net, 
              metric == "H2" & ad.metric == "fl.abund"),
       aes(log(ad.value), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, 
              metric == "H2" & ad.metric == "fl.rich"),
       aes(log(ad.value), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, 
              metric == "H2" & ad.metric == "bb.abund"),
       aes(log(ad.value), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, 
              metric == "H2" & ad.metric == "bb.rich"),
       aes(ad.value, value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, metric == "weighted.NODF"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_net, metric == "web.asymmetry"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_sp, 
              metric == "d" &
                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
                ad.metric == "fl.rich"), 
       aes(ad.value, value, color = year)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(sp ~ phase, scales = "free") +
  ylim(c(0,1))

pxe_sp_mean <- pxe_sp %>%
  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
  group_by(site, year, phase, elev.mean, metric) %>%
  summarize(value = mean(value))

ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value, color = metric)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(pxe_sp, 
              metric == "species.strength",
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
       aes(elev.mean, value, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(sp ~ phase, scales = "free")

>>>>

### Prepare environment
```{r setup, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
#library(lme4)
#library(lmerTest)
library(tidyverse)
library(lubridate)
library(GGally)
```

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      filter(freq > 1) %>% # remove singletons
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species specificity")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality.HL")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

# network level PxE
rare_net_nt_pxe <- function(net, n, iter) {

@@

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      select(site, date, phase, year) %>%
      unique() %>%
      group_by(site, phase, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site, phase, year) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      unite(webID, c(site, phase, year), sep = "_") %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher",
          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(id = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      select(site, date, phase, year) %>%
      unique() %>%
      group_by(site, phase, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site, phase, year) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      unite(webID, c(site, phase, year), sep = "_") %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "id") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

### Colony phase breakpoint analysis
```{r}
castes <- read_csv("./processed_data/network.csv") %>%
  filter(bb.sp != "psit") %>% # We do not want to consider Psithyrus, since these parasitic species do not have a worker caste
  group_by(site, date, year, caste) %>%
  summarize(caste.count = n()) %>%
  spread(caste, caste.count) %>%
  replace(is.na(.), 0) %>%
  mutate(queen = oldqueen + queen,
         reproductive = male + youngqueen,
         q.prop = queen/(queen+worker+reproductive),
         r.prop = reproductive/(queen+worker+reproductive)) %>%
  select(-oldqueen, -youngqueen, -male) %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(site_data, by = c("site")) %>%
  arrange(gdd.cum) %>%
  mutate(yday = yday(date))

breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.5, t2 = 0.15)
breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.5, t2 = 0.15)
breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.5, t2 = 0.15)

breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.15)
breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.15)
breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.15)

@@

### Plot phase breakpoints
```{r}
breaks_summary <- full_join(breaks_gdd_summary, breaks_yday_summary) %>%
  select(site, date, year, elev.class2, yday, gdd.cum, break1, break2, caste, prop) %>%
  gather(time.metric, value, -c(site, date, year, elev.class2, break1, break2, caste, prop))

ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
  geom_point(alpha = 0.5) +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("Julian day") +
  ylab("Proportional abundance")

ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
  geom_point(alpha = 0.5) +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("GDD") +
  ylab("Proportional abundance")

ggplot(breaks_summary, aes(value, prop, color = caste)) +
  geom_point(alpha = 0.25) +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  facet_grid(elev.class2~ time.metric, scales = "free") +
  xlab("Julian day") +
  ylab("Proportional abundance") 
```

@@

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  #semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = "site") %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)
```
### BB and floral turnover ~ elevation
```{r}
bb_range <- net %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

@@

ggplot(filter(fl_survey_elev, sites > 0), 
       aes(reorder(plant.sp, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5) +
  theme_light(14) +
  theme(axis.text.x = element_blank()) +
  xlab("Plant species") +
  ylab("Elevation")

ggplot(filter(fl_net_elev, sites > 0), 
       aes(reorder(plant.sp, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5) +
  theme_light(14) +
  theme(axis.text.x = element_blank()) +
  xlab("Plant species") +
  ylab("Elevation")

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 3, color = "gray60") +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_light(14) +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### BB abundance, diversity, and per-capita resource availability
```{r eval=FALSE, echo=FALSE}
### Per-transect Bombus density
bb_abund <- net %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.abund = n())

### Per-transect Bombus richness
bb_rich <- net %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2,
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.rich = length(unique(bb.sp)))

### Per-transect floral abund
fl_abund <- survey %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.abund = sum(flower.cover),
            log.fl.abund = log(sum(flower.cover)))

### Per-transect floral richness
fl_rich <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.rich = n())

### Join BB density and floral cover
abund_div <- bb_abund %>%
  full_join(bb_rich) %>%
  full_join(fl_abund) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
  full_join(fl_rich) %>%
  #na.omit() %>%
  ungroup %>%
  mutate(fl.per.bb = fl.abund/bb.abund,
         site = factor(site),
         yday = dayofyear,
         week = week(date),
         month = month(date)) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")),
         total.rich = bb.rich + fl.rich,
         web.asymmetry = (bb.rich - fl.rich) / (bb.rich + fl.rich))

abund_div_mean <- abund_div %>%
  select(-dayofyear, -week, -month, -break1.gdd, -break2.gdd) %>%
  gather(metric, value, -c(site, date, gdd.cum, year, elev.class2, elev.mean, phase)) %>%
  group_by(site, year, elev.class2, elev.mean, metric) %>%
  summarize(value = mean(value, na.rm = TRUE))

@@

ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  facet_grid(year~., scales = "free") +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(abund_div, aes(elev.mean, fl.per.bb)) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
  geom_point(alpha = 1) +
  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  facet_wrap(~year) +
  theme_light()

lm1 <- lm(value ~ elev.mean * year, data = filter(abund_div_mean, metric == "fl.per.bb"))
summary(lm1)

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, value)) +
  geom_point(alpha = 1) +
  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  geom_smooth() +
  ylab("fl-per-bb") +
  xlab("Elevation") +
  theme_light()

@@

ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  ylab("log(fl-per-bb)") +
  xlab("GDD") +
  theme_light()

@@

ggplot(filter(abund_div_mean, metric == "web.asymmetry"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ ., scales = "free")

### Web asymmetry ~ gdd
ggplot(abund_div, aes(gdd.cum, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~.)

### Web asymmetry ~ yday
ggplot(abund_div, aes(yday, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~.)

@@

### Web asymmetry ~ Per-capita floral resource abundance
ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ .)
```

### GAM surface plots of abundance and diversity metrics
#### HGAM elevation x time x year (I would really like someone -- maybe my future self -- to check that I'm doing this right)
##### Floral abundance
```{r}
fl_abund_yday <- gam(fl.abund ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    family = "tw",
                    method = "REML")

fl_abund_yday.vis <- getViz(fl_abund_yday)
check.gamViz(fl_abund_yday.vis)
summary(fl_abund_yday.vis)
plot(sm(fl_abund_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance 2010 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(fl_abund_yday.vis, 2)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance 2011 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(fl_abund_yday.vis, 3)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance 2012 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

fl_abund_gdd <- gam(fl.abund ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    family = "tw",
                    method = "REML")

fl_abund_gdd.vis <- getViz(fl_abund_gdd)
check.gamViz(fl_abund_gdd.vis)
summary(fl_abund_gdd.vis)
plot(sm(fl_abund_gdd.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance 2010 (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(fl_abund_gdd.vis, 2)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance 2011 (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(fl_abund_gdd.vis, 3)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance 2012 (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")
```

##### BB abundance
```{r}
bb_abund_yday <- gam(bb.abund ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    family = "scat",
                    method = "REML")

bb_abund_yday.vis <- getViz(bb_abund_yday)
check.gamViz(bb_abund_yday.vis)
summary(bb_abund_yday.vis)
plot(sm(bb_abund_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("bb abundance 2010 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(bb_abund_yday.vis, 2)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("bb abundance 2011 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(bb_abund_yday.vis, 3)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("bb abundance 2012 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

bb_abund_gdd <- gam(bb.abund ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")
